SymPy includes a basic version of the module for dealing with categories ---
abstract mathematical objects representing classes of structures as classes of
objects (points) and morphisms (arrows) between the objects.  This version of
the module was designed with the following two goals in mind:

\begin{enumerate}
\item automatic typesetting of diagrams given by a collection of
  objects and of morphisms between them, and
\item specification and (semi-)automatic derivation of properties
  using commutative diagrams.
\end{enumerate}

At of version 1.0, SymPy only implements the first goal, while a (very partially
working) draft of implementation of the second goal is available
at~\cite{ct4commutativity}.

In order to achieve the two goals, the module \texttt{categories} defines
several classes representing some of the essential concepts: objects, morphisms,
categories, and diagrams.  In category theory, the inner structure of objects is
often discarded in the favor of studying the properties of morphisms, so the
class \texttt{Object} is essentially a synonym of the class \texttt{Symbol}.
There are several morphism classes which do not have a particular internal
structure either, though an exception is \texttt{CompositeMorphism}, which
essentially stores a list of morphisms.

To capture the properties of morphisms, the class \texttt{Diagram} is expected
to be used.  This class stores a family of morphisms, the corresponding source
and target objects, and, possibly, some properties of the morphisms.  Generally,
no restrictions are imposed on what the properties may be --- for example, one
might use strings of the form ``forall'', ``exists'', ``unique'', etc.
Furthermore, the morphisms of a diagram are grouped into \textit{premises} and
\textit{conclusions}, in order to be able to represent logical implications of
the form ``for a collection of morphisms $P$ with properties $p:P\to \Omega$ (the
premises), there exists a collection of morphisms $C$ with properties $c:C\to
\Omega$ (the conclusions),'' where $\Omega$ is the universal collection of
properties.  Finally, the class \texttt{Category} includes a collection of
diagrams which are deemed commutative and which therefore define the properties
of this category.

Automatic typesetting of diagrams takes a \texttt{Diagram} and produces \LaTeX{}
code using the \texttt{Xy-pic} package.  Typesetting is done in two stages:
layout and generation of \texttt{Xy-pic} code.  The layout stage is taken care
of by the class \texttt{DiagramGrid}, which takes a \texttt{Diagram} and lays out
the objects in a grid, trying to reduce the average length of the arrows in the
final picture.  By default, \texttt{DiagramGrid} uses a series of triangle-based
heuristics to produce a rectangular grid.  A linear layout can also be imposed.
Furthermore, groups of objects can be given; in this case, the groups will be
treated as atomic cells, and the member objects will be typeset independently of
the other objects.

The second phase of diagram typesetting consists of actually drawing the picture
and is carried out by the class \texttt{XypicDiagramDrawer}.  An example of a
diagram automatically typeset by \texttt{DiagramgGrid} and
\texttt{XypicDiagramDrawer} in given in Figure~\ref{fig:cat:loops}.
\begin{figure}[h]
  \centerline{
    \xymatrix{
      A \ar[r]_{f} \ar@/^3mm/[rr]^{h_{2}} \ar@(u,l)[]^{l_{A}} \ar@/^3mm/@(l,d)[]^{n_{A}} & B \ar[d]^{g} & D \ar[l]^{k} \ar@/_7mm/[ll]_{h} \ar@/_11mm/[ll]_{h_{1}} \ar@(r,u)[]^{l_{D}} \ar@/^3mm/@(d,r)[]^{n_{D}} \\
      & C \ar@(l,d)[]^{l_{C}} \ar@/^3mm/@(d,r)[]^{n_{C}} &
    }
  }
  \caption{An automatically typeset commutative diagram}
  \label{fig:cat:loops}
\end{figure}

As far as the second main goal of the module is concerned, a (non-working) draft
of an implementation is at~\cite{ct4commutativity}.  The principal idea consists
of automatically deciding whether a diagram is commutative or not, given a
collection of ``axioms'' --- diagrams \textit{known} to be commutative.  The
implementation is based on graph embeddings (injective maps): whenever an
embedding of a commutative diagram into a given diagram is found, one concludes
that the subdiagram is commutative.  Deciding commutativity of the whole diagram
is therefore based (theoretically) on finding a ``cover'' of the target diagram
by embeddings of the axioms.  The na\"{i}ve implementation proved to be
prohibitively slow; a better optimized version is therefore in order, as well as
application of heuristics.

Contributions to automatic inference of commutativity of diagrams are welcome.
The source code (both the one in master and in \texttt{ct4-commutativity}) is
extensively documented.  Even more extensive explanations (including some
literary chatter) are given at~\cite{scolobb}.
