
SymPy supports construction and manipulation of boolean expressions
through the \texttt{sympy.logic} module. SymPy symbols can be used as
propositional variables and also be substituted as \texttt{True}
or \texttt{False}. A good number of manipulation features for boolean
expressions have been implemented in the \texttt{sympy.logic} module.

\subsubsection{Constructing boolean expressions}

A boolean variable can be declared as a SymPy \verb|Symbol|. Python operators
\texttt{\&}, \texttt{\textbar{}} and \texttt{\textasciitilde{}} are overridden
when using SymPy objects to use the SymPy functionality for logical
\texttt{And}, \texttt{Or}, and \texttt{Not}. Other logic functions are also
integrated into SymPy, including \texttt{Xor} and \texttt{Implies}, which are
constructed with \texttt{\^{}} and \texttt{\textgreater{}\textgreater{}},
respectively. The above are just a shorthand, expressions can also be
constructed by directly creating the relevant objects: \verb|And()|,
\verb|Or()|, \verb|Not()|, \verb|Xor()|, \verb|Implies()|, \verb|Nand()|,
\verb|Nor()|, etc.

\begin{verbatim}
>>> from sympy import *
>>> x, y, z = symbols('x y z')
>>> e = (x & y) | z
>>> e.subs({x: True, y: True, z: False})
True
\end{verbatim}

\subsubsection{CNF and DNF}

Any boolean expression can be converted to conjunctive normal form, disjunctive
normal form, and negation normal form. The API also exposes methods to check if
a boolean expression is in any of the above mentioned forms.

\begin{verbatim}
>>> from sympy.logic.boolalg import is_dnf, is_cnf
>>> x, y, z = symbols('x y z')
>>> to_cnf((x & y) | z)
And(Or(x, z), Or(y, z))
>>> to_dnf(x & (y | z))
Or(And(x, y), And(x, z))
>>> is_cnf((x | y) & z)
True
>>> is_dnf((x & y) | z)
True
\end{verbatim}

\subsubsection{Simplification and Equivalence}

The module supports simplification of given boolean expression by making
deductions from the expression. Equivalence of two logical expressions can also
be checked. In the case of equivalence, it is possible to return the mapping of
variables in two expressions so as to represent the same logical behavior.

\begin{verbatim}
>>> from sympy import *
>>> a, b, c, x, y, z = symbols('a b c x y z')
>>> e = a & (~a | ~b) & (a | c)
>>> simplify(e)
And(Not(b), a)
>>> e1 = a & (b | c)
>>> e2 = (x & y) | (x & z)
>>> bool_map(e1, e2)
(And(Or(b, c), a), {a: x, b: y, c: z})
\end{verbatim}

\subsubsection{SAT solving}

The module also supports satisfiability (SAT) checking of a given boolean
expression. If satisfiable, it is possible to return a model for which the
expression is satisfiable. The API also supports returning all possible models.
The SAT solver has a clause learning DPLL algorithm implemented with a watch
literal scheme and VSIDS heuristic\cite{moskewicz2008method}.

\begin{verbatim}
>>> from sympy import *
>>> a, b, c = symbols('a b c')
>>> satisfiable(a & (~a | b) & (~b | c) & ~c)
False
>>> satisfiable(a & (~a | b) & (~b | c) & c)
{a: True, b: True, c: True}
\end{verbatim}
