% Polynomials

SymPy implements a wide variety of algorithms for polynomial manipulation,
which ranges from relatively simple algorithms for doing arithmetics of
polynomials, to advanced methods for factoring multivariate polynomials
into irreducibles, symbolically determining real and complex root isolation
intervals, or computing Gröbner bases.

Polynomial manipulation is useful on its own, but in SymPy, it's mostly
used indirectly as a tool in other areas of the library. In fact, many
mathematical problems in symbolic computing are first expressed using
entities from the symbolic core, preprocessed, and then transformed into
a problem in the polynomial algebra, where generic and efficient algorithms
are used to solve the problem and in the end, solutions to original one
are recovered. For example, this is a common scheme in symbolic integration
or summation algorithms.

SymPy implements dense and sparse polynomial representations. Both
are used in univariate and multivariate cases. Dense representation is
the default for univariate polynomials. For multivariate polynomials,
the choice of representation is based on the application. The most common
case for sparse representation is algorithms for computing Gröbner bases
(Buchberger, F4 and F5), because different monomial orderings can be
expressed easily in this representation. However, algorithms for computing
multivariate GCDs or factorizations, at least those currently implemented
in SymPy, are better expressed when the representation is dense. By dense
multivariate representation we mean a recursively dense representation,
where polynomial $K[x_0, x_1,\dots, x_n]$ is viewed as a polynomial in
$K[x_0][x_1]\ldots[x_n]$. Note that despite this, the coefficient domain
$K$, can be a multivariate polynomial domain as well. Dense recursive
representation in Python gets inefficient when the number of variables
gets high.

\noindent Factorization:
\begin{verbatim}
>>> var("x,y,z,t")
>>> f = 2115*x**4*y + 45*x**3*z**3*t**2 - 45*x**3*t**2 - 423*x*y**4 - \
        47*x*y**3 + 141*x*y*z**3 + 94*x*y*z*t - 9*y**3*z**3*t**2 +    \
        9*y**3*t**2 - y**2*z**3*t**2 + y**2*t**2 + 3*z**6*t**2 +      \
        2*z**4*t**3 - 3*z**3*t**2 - 2*z*t**3
>>> factor(f)
(47*x*y + z**3*t**2 - t**2)*(45*x**3 - 9*y**3 - y**2 + 3*z**3 + 2*z*t)
\end{verbatim}

\noindent Gröbner bases:
\begin{verbatim}
>>> var('x:3')
>>> I = [x0 + 2*x1 + 2*x2 - 1, x0**2 + 2*x1**2 + 2*x2**2 - x0, 2*x0*x1 + 2*x1*x2 - x1]
>>> groebner(I, oder='lex')
GroebnerBasis([
    7*x0 - 420*x2**3 + 158*x2**2 + 8*x2 - 7,
    7*x1 + 210*x2**3 - 79*x2**2 + 3*x2,
    84*x2**4 - 40*x2**3 + x2**2 + x2], x0, x1, x2, domain='ZZ', order='lex')
\end{verbatim}

\noindent Root isolation:
\begin{verbatim}
>>> var('z')
>>> f = 7*z**4 - 19*z**3 + 20*z**2 + 17*z + 20
>>> intervals(f, all=True, eps=0.001)
([],
 [((-425/1024 - 625*I/1024, -1485/3584 - 2185*I/3584), 1),
  ((-425/1024 + 2185*I/3584, -1485/3584 + 625*I/1024), 1),
  ((3175/1792 - 2605*I/1792, 1815/1024 - 10415*I/7168), 1),
  ((3175/1792 + 10415*I/7168, 1815/1024 + 2605*I/1792), 1)])
\end{verbatim}
